---
title: "Scoping Review - SES and OEL Data Cleaning"
format: html
editor: visual
---

##### Load Libraries

```{r}
library(here)
library(rio)
library(tidyverse)
library(janitor)
library(ggthemes)
library(gghighlight)
library(tidyr)
library(readxl)
library(RColorBrewer)
library(knitr)
library(kableExtra)
library(writexl)
```

### Load numerical data

```{r}
here("data", "FullDataSESOEL22024.xlsx")
ndf1 <- read_excel("/Users/daniellejones/Desktop/Scoping Review Coding/Scoping Review - Qualtrics Data Cleaning/data/FullDataSESOEL22024.xlsx")

ndf1 <- ndf1 %>%
  characterize() %>%
  clean_names()
```

## Load choice data

```{r}
here("data", "FullDataSESOELChoice.xlsx")
cdf1 <- read_excel("/Users/daniellejones/Desktop/Scoping Review Coding/Scoping Review - Qualtrics Data Cleaning/data/FullDataSESOELChoice.xlsx")

cdf1 <- cdf1 %>%
  characterize() %>%
  clean_names()
```

## Take away question row - numerical data

```{r}
ndf2 <- ndf1 %>%
  slice(-1, ) 
```

## Take away question row - choice data

```{r}
cdf2 <- cdf1 %>%
  slice(-1, ) 
```

## Creating individual data sets for each question - numerical data

```{r}
nq2 <- ndf2 %>%
  select("row_id", "q1", "q2")
nq3 <- ndf2 %>%
  select("row_id", "q1", "q3", "q3_44_text")
nq4 <- ndf2 %>%
  select("row_id", "q1", "q4")
nq5 <- ndf2 %>%
  select("row_id", "q1", "q5")
nq6 <- ndf2 %>%
  select("row_id", "q1", "q6", "q6_3_text", "q6a", "q6a_6_text")
nq7 <- ndf2 %>%
  select("row_id", "q1", "q7", "q6_3_text")
nq8 <- ndf2 %>%
  select("row_id", "q1", "q8", "q8_5_text", "q8_6_text")
nq9 <- ndf2 %>%
  select("row_id", "q1", "q9", "q9_3_text", "q9_4_text")
nq10 <- ndf2 %>%
  select("row_id", "q1", "q10", "q10_5_text", "q10_west", "q10_midwest", "q10_south", "q10_northeast")
nq11 <- ndf2 %>%
  select("row_id", "q1", "q11", "q11_1_text", "q11_3_text")
nq12 <- ndf2 %>%
  select("row_id", "q1", "q12", "q12_1_text", "q12_2_text", "q12_3_text", "q12_4_text", "q12_5_text", 
         "q12_6_text", "q12_7_text", "q12_8_text")
nq13 <- ndf2 %>%
  select("row_id", "q1", "q13", "q13_13_text", "q13_14_text", "q13_15_text")
nq14 <- ndf2 %>%
  select("row_id", "q1", "q14")
nq15 <- ndf2 %>%
  select("row_id", "q1", "q15", "q15_44", "q15_7_text_45", "q15_8_text", "q15_47", "q15_7_text_48",
         "q15_6_text", "q15_7_text")
nq16 <- ndf2 %>%
  select("row_id", "q1", "q16", "q16_1_text", "q16_2_text", "q16_3_text", "q16_4_text", "q16_5_text",
          "q16_6_text", "q16_7_text", "q16_8_text")
nq17 <- ndf2 %>%
  select("row_id", "q1", "q17", "q17_1_text", "q17_2_text", "q17_3_text", "q17_4_text", "q17_5_text",
          "q17_6_text", "q17_7_text", "q17_8_text")
nq18 <- ndf2 %>%
  select("row_id", "q1", "q18", "q18_6_text", "q18_7_text")
nq19 <- ndf2 %>%
  select("row_id", "q1", "q19")
nq20 <- ndf2 %>%
  select("row_id", "q1", "q20", "q20_6_text")
nq21 <- ndf2 %>%
  select("row_id", "q1", "q21")
nq22 <- ndf2 %>%
  select("row_id", "q1", "q22", "q22_1_text", "q22_2_text", "q22_3_text", "q22_4_text")
nq23 <- ndf2 %>%
  select("row_id", "q1", "q23", "q23_4_text")
nq24 <- ndf2 %>%
  select("row_id", "q1", "q24")
nq25 <- ndf2 %>%
  select("row_id", "q1", "q25")
nq26 <- ndf2 %>%
  select("row_id", "q1", "q26", "q26_b_1_text", "q26_b_2_text", "q26_b_3_text", "q26_b_4_text",
         "q26_b_5_text", "q26_b_6_text", "q26_b_7_text")
nq27 <- ndf2 %>%
  select("row_id", "q1", "q27", "q27_1_text", "q27_6_text")
nq28 <- ndf2 %>%
  select("row_id", "q1", "q28", "q28_1_text", "q28_2_text", "q28_3_text", "q28_4_text", "q28_5_text")
nq29 <- ndf2 %>%
  select("row_id", "q1", "q29", "q29_1_text", "q29_4_text")
nq30 <- ndf2 %>%
  select("row_id", "q1", "q30", "q30_11_text", "q30_12_text", "q30_13_text")
nq31 <- ndf2 %>%
  select("row_id", "q1", "q31", "q31_6_text")

```

## Creating individual data sets for each question - choice data

```{r}
cq2 <- cdf2 %>%
  select("row_id", "q1", "q2")
cq3 <- cdf2 %>%
  select("row_id", "q1", "q3", "q3_44_text")
cq4 <- cdf2 %>%
  select("row_id", "q1", "q4")
cq5 <- cdf2 %>%
  select("row_id", "q1", "q5")
cq6 <- cdf2 %>%
  select("row_id", "q1", "q6", "q6_3_text", "q6a", "q6a_6_text")
cq7 <- cdf2 %>%
  select("row_id", "q1", "q7", "q6_3_text")
cq8 <- cdf2 %>%
  select("row_id", "q1", "q8", "q8_5_text", "q8_6_text")
cq9 <- cdf2 %>%
  select("row_id", "q1", "q9", "q9_3_text", "q9_4_text")
cq10 <- cdf2 %>%
  select("row_id", "q1", "q10", "q10_5_text", "q10_west", "q10_midwest", "q10_south", "q10_northeast")
cq11 <- cdf2 %>%
  select("row_id", "q1", "q11", "q11_1_text", "q11_3_text")
cq12 <- cdf2 %>%
  select("row_id", "q1", "q12", "q12_1_text", "q12_2_text", "q12_3_text", "q12_4_text", "q12_5_text", 
         "q12_6_text", "q12_7_text", "q12_8_text")
cq13 <- cdf2 %>%
  select("row_id", "q1", "q13", "q13_13_text", "q13_14_text", "q13_15_text")
cq14 <- cdf2 %>%
  select("row_id", "q1", "q14")
cq15 <- cdf2 %>%
  select("row_id", "q1", "q15", "q15_44", "q15_7_text_45", "q15_8_text", "q15_47", "q15_7_text_48",
         "q15_6_text", "q15_7_text")
cq16 <- cdf2 %>%
  select("row_id", "q1", "q16", "q16_1_text", "q16_2_text", "q16_3_text", "q16_4_text", "q16_5_text",
          "q16_6_text", "q16_7_text", "q16_8_text")
cq17 <- cdf2 %>%
  select("row_id", "q1", "q17", "q17_1_text", "q17_2_text", "q17_3_text", "q17_4_text", "q17_5_text",
          "q17_6_text", "q17_7_text", "q17_8_text")
cq18 <- cdf2 %>%
  select("row_id", "q1", "q18", "q18_6_text", "q18_7_text")
cq19 <- cdf2 %>%
  select("row_id", "q1", "q19")
cq20 <- cdf2 %>%
  select("row_id", "q1", "q20", "q20_6_text")
cq21 <- cdf2 %>%
  select("row_id", "q1", "q21")
cq22 <- cdf2 %>%
  select("row_id", "q1", "q22", "q22_1_text", "q22_2_text", "q22_3_text", "q22_4_text")
cq23 <- cdf2 %>%
  select("row_id", "q1", "q23", "q23_4_text")
cq24 <- cdf2 %>%
  select("row_id", "q1", "q24")
cq25 <- cdf2 %>%
  select("row_id", "q1", "q25")
cq26 <- cdf2 %>%
  select("row_id", "q1", "q26", "q26_b_1_text", "q26_b_2_text", "q26_b_3_text", "q26_b_4_text",
         "q26_b_5_text", "q26_b_6_text", "q26_b_7_text")
cq27 <- cdf2 %>%
  select("row_id", "q1", "q27", "q27_1_text", "q27_6_text")
cq28 <- cdf2 %>%
  select("row_id", "q1", "q28", "q28_1_text", "q28_2_text", "q28_3_text", "q28_4_text", "q28_5_text")
cq29 <- cdf2 %>%
  select("row_id", "q1", "q29", "q29_1_text", "q29_4_text")
cq30 <- cdf2 %>%
  select("row_id", "q1", "q30", "q30_11_text", "q30_12_text", "q30_13_text")
cq31 <- cdf2 %>%
  select("row_id", "q1", "q31", "q31_6_text")

```

## Q2 & Q3 - Journals and Publication Years

##### Line Graph of Journal Publication Year

```{r}

cq2_counts <- cq2 %>%
  rename(year = q2) %>%
  mutate(year = as.integer(year))

cq2_year_counts <- cq2_counts %>%
  group_by(year) %>%
  mutate(year_count = n()) %>%
  ungroup()

year_plot <-
  ggplot(cq2_year_counts, aes(x = year, y = year_count)) +
  geom_point() + 
  geom_line() +  
  scale_x_continuous(breaks = seq(1980, 2023, by = 2)) +
  scale_y_continuous(breaks = seq(0, 8, by = 1), limits = c(0, NA)) +
  labs(x = "Year", y = "Number of Articles", title = "Publication of SES and OEL Articles Over the Years") +
  theme(
    panel.background = element_blank())
print(year_plot)

```

##### Table of Publication Frequency by Year

```{r}
year_counts_table_df <- cq2_year_counts %>%
  distinct(year, .keep_all = TRUE) %>%
  select(year, year_count) %>%
  arrange(desc(year)) %>%
  rename("Year" = "year", "Article Count" = "year_count")

year_counts_prettytable <- knitr::kable(year_counts_table_df, "html") %>%
  kableExtra::kable_styling(full_width = FALSE)

print(year_counts_prettytable)
```

##### Table of all publications

```{r}
cq3_counts <- cq3 %>%
  mutate(q3 = as.character(q3)) %>%
  group_by(q3) %>%
  summarise(q3_count = n()) %>%
  arrange(desc(q3_count)) %>%
  rename("Journal" = "q3", "Article Count" = "q3_count")

publications_all_prettytable <- knitr::kable(cq3_counts, "html") %>%
  kableExtra::kable_styling(full_width = FALSE)

print(publications_all_prettytable)

```

##### Top Journals for Publication Data Prep

```{r}
nq3_counts <- nq3 %>%
  rename(journal = q3) %>%
  mutate(journal = as.integer(journal)) %>% 
  group_by(journal) %>%
  mutate(journal = as.character(journal)) %>%
  mutate(journal_count = n()) %>%
  summarise(journal_count =n()) 



```

##### Top Journals for Publication Plot

```{r}
top_11_journals <- nq3_counts %>%
   mutate(journal = factor(journal, 
                            levels = c("10", 
                                       "6", 
                                       "15", 
                                       "33", 
                                       "14", 
                                       "24", 
                                       "35", 
                                       "12", 
                                       "4", 
                                       "18", 
                                       "43"),
                            labels = c("Developmental Psychology", 
                                       "Child Development", 
                                       "Early Education and Development", 
                                       "Journal of Speech, Language, and Hearing Research", 
                                       "Early Childhood Research Quarterly",
                                       "Journal of Child Language",
                                       "Language, Speech, and Hearing Services in Schools",
                                       "Early Child Development and Care",
                                       "Applied Pscyholinguistics",
                                       "Infant and Child Development",
                                       "Topics in Early Childhood Special Education"
                                       ))) %>%
  arrange(desc(journal_count)) %>%
  head(11)

journal_plot <- ggplot(top_11_journals, aes(x = reorder(journal, journal_count), y = journal_count, fill = journal)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  geom_text(aes(label = journal_count), vjust = .5, color = "black", size = 3, hjust = -1) +  
  labs(x = "", y = "Number of Articles from 1980 - 2023", title = "Journals with Highest Frequency of Publication of Articles About SES and OEL") +
  scale_y_continuous(breaks = seq(0, 10, by = 1)) +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        plot.title = element_text(hjust = 10),
         axis.ticks.x = element_line(color = "black")) +
  coord_flip()

print(journal_plot)
```

## Q6 - Study Design

##### Separate Quant/Mixed from Quant Description

##### Create frequency count

```{r}
q6_frequency <- nq6 %>%
  count(q6)
```

##### Study Design Bar Chart

```{r}
methods_bar <- 
  ggplot(q6_frequency, aes(q6, n, fill = q6)) +
  geom_bar(stat= "identity", show.legend = FALSE) +
  geom_text(aes(label = n), vjust = -0.5, size = 3) +
  labs(title = "Methods Used to Study SES and OEL", 
       y = "Number of Studies",
       x ="") +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(breaks = c(0,10,20,30,40,50, 60, 70, 80, 90, 100),
                       limits = c(0,100),
                       expand = c(0,0))+
  scale_x_discrete(labels = c("1" = "Quantitative", 
                              "2" = "Mixed Methods",
                              "3" = "Other/Unsure")) +
  theme(
  panel.background   = element_blank(),
  axis.line.x        = element_line(color = "black", linewidth = 0.5),
  axis.line.y        = element_line(color = "black", linewidth = 0.5),
)

print(methods_bar)
```

##### Study design percentage

```{r}
q6_total_count <- sum(q6_frequency$n)
q6_percent <- q6_frequency %>%
  mutate(percentage = (n / total_count) * 100)

print(q6_percent)
```

## Q6a - Quantative Methods Used

##### Selecting 6a column, pivoting long, then pivoting wide, summarizing, percentages

```{r}
q6a <- nq6 %>%
  select(row_id, q6a) %>%
  separate_rows(q6a, sep = ",") 

q6a_wide <- q6a %>%
  pivot_wider(names_from = q6a, values_from = q6a, values_fn = length, values_fill = 0)

q6a_wide_totals <- q6a_wide %>%
  adorn_totals("row")

q6a_wide_totals <- q6a_wide_totals %>%
  rename("Descriptive-Observational" = "1",
         "Descriptive-Correlational" = "2",
         "Quasi-Experimental" = "3")

q6a_df <- data.frame(
  `Descriptive-Observational` = 46,
  `Descriptive-Correlational` = 57,
  `Quasi-Experimental` = 6,
  `Single-Subject` = 0)

print(q6a_df)

```

## Q7 - Data Collection Methods

```{r}
q7_frequency <- nq7 %>%
  count(q7)

q7_total_count <- sum(q7_frequency$n)
q7_percent <- q7_frequency %>%
  mutate(percentage = round((n / q7_total_count) * 100, digits = 1))

print(q7_percent)
```

## Q8 - Setting/Context

```{r}

```

## Q9 - Urban/Rural

##### Recode variables to 6 if both urban and rural populations were studied

```{r}
nq9_6 <- nq9 %>%
  mutate(q9 = ifelse(lengths(strsplit(q9, ",")) == 2, "6", q9))

q9_6_frequency <- nq9_6 %>%
  count(q9)
```

##### Urban/Rural percentage

```{r}
q9_6_total_count <- sum(q9_6_frequency$n)
q9_6_percent <- q9_6_frequency %>%
  mutate(percentage = round((n / q9_6_total_count) * 100, digits = 1))

print(q9_6_percent)
```

##### Bar graph for urban/rural with both option

```{r}
urban_rural_bar <- 
  ggplot(q9_6_frequency, aes(q9, n, fill = q9)) +
  geom_bar(stat= "identity", show.legend = FALSE) +
  geom_text(aes(label = n), vjust = -0.5, size = 3) +
  labs(title = "Urban, Rural, Both", 
       y = "Number of Studies",
       x ="") +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(breaks = c(0,10,20,30,40,50, 60, 70),
                       limits = c(0,70),
                       expand = c(0,0))+
  scale_x_discrete(labels = c("1" = "Urban", 
                              "2" = "Rural",
                              "3" = "Other",
                              "4" = "Unclear",
                              "5" = "Unreported",
                              "6" =" Both Urban & Rural")) +
  theme(
  panel.background   = element_blank(),
  axis.line.x        = element_line(color = "black", linewidth = 0.5),
  axis.line.y        = element_line(color = "black", linewidth = 0.5),
)

print(urban_rural_bar)
```

##### Pivot table longer so each column only has one value

```{r, eval=FALSE}
nq9_long <- nq9 %>%
  select(row_id, q1, q9) %>%
  separate_rows(q9, sep = ",") 

```

## Q10 - Region of the United States

## Q11 - Use and Definition of SES

```{r}

```

## Q12 - Labels

```{r}

```

## Q13 - Measures used to determine SES

```{r}

```

## Q14 - Level of SES Studied

```{r}

```

## Q15 - 5 Domains

## Q16 - Labels

## Q17 - Definition of Labels

## Q18 - Data collection tool used to measure OEL

## Q19 - Analytical Approaches

## Q20 - How SES was used in analysis

##### Pivoting data long then wide to capture columns with multiple data pieces

```{r}
q20_long <- nq20 %>%
  select(row_id, q20) %>%
  separate_rows(q20, sep = ",") 

q20_wide <- q20_long %>%
  pivot_wider(names_from = q20, values_from = q20, values_fn = length) %>%
  select(row_id, `1`, `2`, `3`, `4`, `5`) %>%
  mutate(across(`1`:`5`, ~replace(., is.na(.), 0)))


```

##### Isolating columns that used SES in more than one way

```{r}
q20_multiple_values <- nq20 %>%
  filter(rowSums(across(where(is.character), ~grepl(",", ., fixed = TRUE))) > 0)

q20_multiple_values_with_labels <- q20_multiple_values %>%
  mutate(label = case_when(
    q20 == "1,2" ~ "inclusion and control",
    q20 == "1,4" ~ "inclusion and moderator",
    q20 == "1,5" ~ "inclusion and independent variable",
    q20 == "2,5" ~ "control and independent")
  )

q20_multiple_count <- q20_multiple_values_with_labels %>%
  count(q20)

print(q20_multiple_count)
```

##### Isolating columns that used SES in one way

```{r}
q20_single_value <- nq20 %>%
  filter(!str_detect(q20, ","))

q20_single_count <- q20_single_value %>%
  count(q20)

print(q20_single_count)
```

##### Combining data sets Q20

```{r}
q20_combined_data <- bind_rows(q20_single_count, q20_multiple_count)
```

##### Bar chart for Q20

```{r}
ggplot(q20_combined_data, aes(x = q20, y = n)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Bar Chart of q20 Counts",
       x = "q20 Categories",
       y = "Count") +
  scale_x_discrete(labels = c("1" = "Inclusion/Exclusion", 
                              "2" = "Covariate/Control", 
                              "3" = "Mediator", 
                              "4" = "Moderator", 
                              "5" = "Independent Variable", 
                              "6" = "Unclear",
                              "1,2" = "Inclusion & Control", 
                              "1,4" = "Inclusion and Moderator", 
                              "1,5" = "Inclusion and Ind. Variable", 
                              "2,5" = "Control and Ind. Variable")) +
  geom_text(aes(label = n), vjust = -0.5, size = 3) +
  labs(title = "Bar Chart of q20 Counts",
       x = "q20 Categories",
       y = "Count") +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Q22 - Inclusion & Exclusion Criteria

```{r}

```

## Q23 - Enrollment in Head Start

##### Head Start Bar Chart

```{r}
q23_frequency <- nq23 %>%
  count(q23) 

headstart_bar <- 
  ggplot(q23_frequency, aes(q23, n, fill = q23)) +
  geom_bar(stat= "identity", show.legend = FALSE) +
  geom_text(aes(label = n), vjust = -0.5, size = 3) +
  labs(title = "Participant Enrollment in Head Start", 
       y = "Number of Studies",
       x ="") +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(breaks = c(0,10,20,30,40,50, 60),
                       limits = c(0,60),
                       expand = c(0,0))+
  scale_x_discrete(labels = c("1" = "Yes - All participants", 
                              "2" = "Some participants - not all",
                              "3" = "No",
                              "4" = "Unclear")) +
  theme(
  panel.background   = element_blank(),
  axis.line.x        = element_line(color = "black", linewidth = 0.5),
  axis.line.y        = element_line(color = "black", linewidth = 0.5),
)

print(headstart_bar)
```

## Q26 - Sex/Gender

##### Reporting of sex/gender

```{r}
q26_frequency <- nq26 %>%
  count(q26) 

gender_report_bar <- 
  ggplot(q26_frequency, aes(q26, n, fill = q26)) +
  geom_bar(stat= "identity", show.legend = FALSE) +
  geom_text(aes(label = n), vjust = -0.5, size = 3) +
  labs(title = "Studies Reporting of Gender", 
       subtitle = "No studies reported sex separate from gender",
       y = "Number of Studies",
       x ="") +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(breaks = c(0,10,20,30,40,50, 60, 70, 80),
                       limits = c(0,80),
                       expand = c(0,0))+
  scale_x_discrete(labels = c("1" = "Authors Described Sex/Gender", 
                              "2" = "Authors Did Not Describe Sex/Gender",
                              "3" = "Author Reported Sex Separate from Gender")) +
  theme(
  panel.background   = element_blank(),
  axis.line.x        = element_line(color = "black", linewidth = 0.5),
  axis.line.y        = element_line(color = "black", linewidth = 0.5),
)

print(gender_report_bar)
```

## Q29 - Disability Status

##### Creating Bar Graph for Disability Status

```{r}
desired_order <- c("1", "2", "3", "4")

q29_frequency <- nq29 %>%
  count(q29) 

q29_frequency <- q29_frequency %>%
  mutate(q29 = fct_relevel(q29, desired_order)) %>%
  mutate(bar_id = as.factor(row_number()))

dis_status_bar <- 
  ggplot(q29_frequency, aes(q29, n, fill = q29)) +
  geom_bar(stat= "identity", show.legend = FALSE) +
  geom_text(aes(label = n), vjust = -0.5, size = 3) +
  labs(title = "Reporting of Child Disability Status in Articles About SES and OEL", 
       y = "",
       x ="") +
  scale_fill_brewer(palette = "Set1") +
  scale_x_discrete(labels = c("1" = "Yes - Specified", 
                              "2" = "Yes - Not Specified",
                              "3" = "No",
                              "4" = "Unclear")) +
  theme(
  panel.background   = element_blank(),
  axis.line.x        = element_line(color = "black", linewidth = 0.5),
                    axis.ticks.y       = element_blank(),  # Remove y-axis ticks
    axis.text.y        = element_blank(),  
    axis.line.y        = element_blank())


print(dis_status_bar)

```

## Q31 - Immigrant Status

```{r}
q31_count_5s <- sum(nq31$q31 == 5)

# Calculate the percentage of 5s
q31_percentage_5s <- (q31_count_5s / nrow(nq31)) * 100

# Print the count and percentage
cat("Number of 5s:", q31_count_5s, "\n")
cat("Percentage of 5s:", q31_percentage_5s, "%\n")


##### 92% of articles did not specify immigration status
```
